---
title: "Natuurindicatoren"
author: "Yasmine Verzelen"
date: "24 oktober 2017"
output: html_document
---

In dit script willen we de data van waarnemingen.be van GBIF halen, voor niet-inheemse planten en dieren.

Load packages
```{r}
library(rgbif)
library(curl)
library(ggplot2)
#library(INBOtheme)
```

Number of records in Waarnemingen.be datasets
```{r}
'Plants'
occ_count(datasetKey='7f5e4129-0717-428e-876a-464fbd5d9a47')
'Animals'
occ_count(datasetKey='9a0b66df-7535-4f28-9f4e-5bc11b8b096c')

```
Get the data from the datasets
```{r}
'Animals'
WnmBe_Animals <- occ_download_get(key = "0006839-171025204727650", overwrite = TRUE) %>% 
    occ_download_import(WnmBe_Animals_download, na.strings = c("", NA))
WnmBe_Animals$species <- as.character(WnmBe_Animals$species)

'Plants'
WnmBe_Plants <- occ_download_get(key = "0006843-171025204727650", overwrite = TRUE) %>% 
    occ_download_import(WnmBe_Plants_download, na.strings = c("", NA))

```

Merge the datasets
```{r}
WnmBe_All <- rbind(WnmBe_Animals, WnmBe_Plants)
```
Calculate the first occurence per species
```{r}
WnmBe_All <- subset(WnmBe_All, !is.na(species))
species <- unique(WnmBe_All$species)
species_list <- data.frame()
for(s in species){
  temp <- subset(WnmBe_All, species == s)
  temp2 <- data.frame(s)
  temp2$firstyear <- min(temp$year)
  temp2$kingdom <- unique(temp$kingdom)
  temp2$phylum <- unique(temp$phylum)
  temp2$class <- unique(temp$class)
  species_list <- rbind(species_list, temp2)
}

remove(temp)
remove(temp2)

```

Calculate the cumulative number of invasive species and display in a graph
```{r}

firstrecord <- min(species_list$firstyear)
newestrecord <- max(unique(WnmBe_All$year))
tijdspanne <- (newestrecord-firstrecord)+1

new_species <- data.frame()
temp2 <- data.frame(1)
t <- firstrecord

for(i in 1:tijdspanne){
  temp <- subset(species_list, firstyear == t)
  temp2$n <- nrow(temp)
  temp2$t <- t
  new_species <- rbind(new_species, temp2)
  t <- t + 1
}

new_species$X1 <- NULL
new_species$cumul <- cumsum(new_species$n)

plot <- ggplot(new_species, aes(x = t, y=cumul)) +
  geom_line(colour="red") +
  xlab("Year") + ylab("Number") + ggtitle("Cumulative number of invasive species") +
  scale_x_continuous(breaks = seq(1800, 2017, 10))+
  scale_y_continuous(breaks = seq(0,1900, 500))
  
print(plot)

```

Calculate the cumulative number of invasive species per kingdom and total and display in a graph
```{r}
firstrecord <- min(species_list$firstyear)
newestrecord <- max(unique(WnmBe_All$year))
tijdspanne <- (newestrecord-firstrecord)+1

new_species_k <- data.frame()
temp2 <- data.frame(1)
temp3 <- data.frame(1)
t <- firstrecord

kingdoms <- unique(species_list$kingdom)
for(k in kingdoms){
  temp <- subset(species_list, kingdom == k)
  for(i in 1:tijdspanne){
    temp2 <- subset(temp, firstyear == t)
    temp3$n <- nrow(temp2)
    temp3$t <- t
    temp3$kingdom <- k
    new_species_k <- rbind(new_species_k, temp3)
    t <- t + 1
  }
  t <- firstrecord
}

new_species_k$X1 <- NULL
new_species_k2 <-  transform(new_species_k,cumul = ave(n , kingdom, FUN = cumsum))

print(plot <- ggplot(new_species_k2, aes(x = t, y=cumul, colour = kingdom))+
  scale_x_continuous(breaks = seq(1800, 2017, 50))+
  scale_y_continuous(breaks = seq(0,1500, 500))+
  geom_line()+ 
  xlab("Year") + ylab("Number") + ggtitle("Cumulative number of invasive species per kingdom")+
  theme(panel.background = element_rect(fill = "aliceblue"))+
  theme(panel.grid.major = element_line(colour = "dark grey"))+
  theme(panel.grid.minor = element_line(colour = "light grey"))+
  theme(legend.title=element_blank())+
  theme(legend.background = element_rect(fill = "aliceblue")))

```

Make a graph of the different classes of invasive animals
```{r}

species_list_a <- subset(species_list, kingdom =="Animalia")

firstrecord <- min(species_list_a$firstyear)
newestrecord <- max(unique(WnmBe_All$year))
tijdspanne <- (newestrecord-firstrecord)+1

new_species_c <- data.frame()
temp2 <- data.frame(1)
temp3 <- data.frame(1)
t <- firstrecord

classes <- unique(species_list_a$class)
for(c in classes){
  temp <- subset(species_list_a, class == c)
  for(i in 1:tijdspanne){
    temp2 <- subset(temp, firstyear == t)
    temp3$n <- nrow(temp2)
    temp3$t <- t
    temp3$class <- c
    new_species_c <- rbind(new_species_c, temp3)
    t <- t + 1
  }
  t <- firstrecord
}

new_species_c$X1 <- NULL
new_species_c2 <-  transform(new_species_c,cumul = ave(n , class, FUN = cumsum))

print(plot <- ggplot(new_species_c2, aes(x=t, y=cumul))+
  geom_bar(position="dodge",stat="identity")+
  scale_x_continuous(breaks = seq(1800, 2017, 50))+
  scale_y_continuous(breaks = seq(0,1500, 500))+
  xlab("Year") + ylab("Number") + ggtitle("Cumulative number of invasive animal species per class")+
  theme(panel.background = element_rect(fill = "aliceblue"))+
  theme(panel.grid.major = element_line(colour = "dark grey"))+
  theme(panel.grid.minor = element_line(colour = "light grey"))+
  theme(legend.title=element_blank())+
  theme(legend.background = element_rect(fill = "aliceblue"))+
    
  facet_wrap(~class, nrow=6))
```

Make a graph of the different classes of invasive plants
```{r}
species_list_p <- subset(species_list, kingdom =="Plantae")

firstrecord <- min(species_list_p$firstyear)
newestrecord <- max(unique(WnmBe_All$year))
tijdspanne <- (newestrecord-firstrecord)+1

new_species_c <- data.frame()
temp2 <- data.frame(1)
temp3 <- data.frame(1)
t <- firstrecord

classes <- unique(species_list_p$class)
for(c in classes){
  temp <- subset(species_list_p, class == c)
  for(i in 1:tijdspanne){
    temp2 <- subset(temp, firstyear == t)
    temp3$n <- nrow(temp2)
    temp3$t <- t
    temp3$class <- c
    new_species_c <- rbind(new_species_c, temp3)
    t <- t + 1
  }
  t <- firstrecord
}

new_species_c$X1 <- NULL
new_species_c2 <-  transform(new_species_c,cumul = ave(n , class, FUN = cumsum))

print(plot <- ggplot(new_species_c2, aes(x=t, y=cumul))+
  geom_bar(position="dodge",stat="identity")+
  scale_x_continuous(breaks = seq(1800, 2017, 50))+
  scale_y_continuous(breaks = seq(0,1500, 500))+
  xlab("Year") + ylab("Number") + ggtitle("Cumulative number of invasive plant species per class")+
  theme(panel.background = element_rect(fill = "aliceblue"))+
  theme(panel.grid.major = element_line(colour = "dark grey"))+
  theme(panel.grid.minor = element_line(colour = "light grey"))+
  theme(legend.title=element_blank())+
  theme(legend.background = element_rect(fill = "aliceblue"))+
  facet_wrap(~class, nrow=6))
```


Gezien er hiervoor gewerkt werd met het eerste jaar van waarneming, maar er soms meerdere waarnemingen waren in dat eerste jaar, kunnen we door de procedure hierboven niet weten of de bijhorende geografische coördinaten wel kloppen. Vandaar eerst onderstaande analyse waardoor we die eerste datum wel weten.
```{r}
WnmBe_first <- WnmBe_All[WnmBe_All$eventdate == ave(WnmBe_All$eventdate, WnmBe_All$species, FUN=min), ]
```

```{r}
library(plyr)
WnmBe_first <- rename(WnmBe_first, c("species"="s"))
temp4 <- merge(species_list, WnmBe_first, all.y=FALSE)
WnmBe_Points <- temp4[temp4$gbifid == ave(temp4$gbifid, temp4$s, FUN=min), ]
```

Kaart maken van de eerste occurences van alle soorten in Vlaanderen.
Hier zijn er wat problemen, oa. met de versie van R waardoor het package niet werkt... Ik zoek naar een betere package met map van Vlaanderen.
```{r}
#install.packages("BelgiumMaps.OpenStreetMap", repos = "http://www.datatailor.be/rcube", type = "source")
```

In plaats van fancy te starten met mooie kaartjes, beginnen we op een eenvoudige manier. 
```{r}
library(rworldmap)

map <- getMap(resolution = "low")
plot(map, xlim = c(2.54, 6), ylim = c(50.70, 51.51), asp = 1)

points(WnmBe_Points$decimallongitude, WnmBe_Points$decimallatitude, col = "red", cex = .6)
```

Grafieken van het aantal eerste waarnemingen per jaar
```{r}
library(plyr)
number1 <- rle(sort(species_list$firstyear ) )
species_list$number1 <- number1[[1]][ match(species_list$firstyear , number1[[2]] ) ]
species_list

print(plot <- ggplot(species_list, aes(x = firstyear, y= number1)) +
  geom_point(colour="black", fill="black")+
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive species per year") +
  scale_x_continuous(breaks = seq(1800, 2017, 10))+
  scale_y_continuous(breaks = seq(0,500, 50)))
```

Duidelijkere grafiek indien vanaf 1940
```{r}
species_list2 <- subset(species_list, firstyear > 1940)

print(plot <- ggplot(species_list2, aes(x = firstyear, y=number1)) +
  geom_point(colour="black", fill="black", size = 4) +
  geom_line(colour= "grey")+
    stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(breaks = seq(0,500, 50)))

```

Make a graph of the number of invasive animal species
```{r}
library(scales)
species_list_2a <- subset(species_list2, kingdom =="Animalia", select = -c(number1))

number2 <- rle(sort(species_list_2a$firstyear))
species_list_2a$number2 <- number2[[1]][ match(species_list_2a$firstyear , number2[[2]] ) ]
species_list_2a

print(plot <- ggplot(species_list_2a, aes(x = firstyear, y=number2)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive animal species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(limit=c(0,NA),oob=squish,breaks = seq(0,NA, 1)))

```

Make a graph of the number of invasive plant species
```{r}
species_list_2p <- subset(species_list2, kingdom =="Plantae", select = -c(number1))

number2 <- rle(sort(species_list_2p$firstyear))
species_list_2p$number2 <- number2[[1]][ match(species_list_2p$firstyear , number2[[2]] ) ]
species_list_2p

print(plot <- ggplot(species_list_2p, aes(x = firstyear, y=number2)) +
  geom_point(colour="black", fill="black", size = 4, position = "dodge",stat="identity") +
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive plant species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(limit=c(0,NA),oob=squish, breaks = seq(0,NA,10)))

```

Combine the three graphs above into one figure:
1. Set up the multiplot function
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=2, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(c(1,2,1,3), nrow = 2)
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

2. Make the figure
```{r}
a1 <- ggplot(species_list2, aes(x = firstyear, y=number1)) +
  geom_point(colour="black", fill="black", size = 4) +
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(breaks = seq(0,500, 50))

a2 <- ggplot(species_list_2a, aes(x = firstyear, y=number2)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive animal species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(limit=c(0,NA),oob=squish)

a3 <- ggplot(species_list_2p, aes(x = firstyear, y=number2)) +
  geom_point(colour="black", fill="black", size = 4, position = "dodge",stat="identity") +
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive plant species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(limit=c(0,NA),oob=squish)


multiplot(a1, a2, a3)

png(filename = "First record rate of invasive animal and plant species per year.png", width = 1000, height = 900, res=80)
multiplot(a1, a2, a3)
dev.off()

```


Make a graph of the number of invasive animal species per class
```{r}
library(dplyr)
library(plyr)

temp5 <- ddply(species_list_2a, .(firstyear, class), nrow)
species_list_3a <- merge(species_list_2a, temp5, by=c("firstyear","class"))

species_list_3a <- subset(species_list_3a, select = -c(number2))
names(species_list_3a)[names(species_list_3a)=="V1"] <- "number3"

print(plot <- ggplot(species_list_3a, aes(x = firstyear, y=number3)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  facet_wrap(~class, nrow=6, scales = "free_y")+
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("Number") + ggtitle("Number of invasive animal species per year per class") +
  scale_y_continuous(limit=c(0,NA), oob=squish))

```

Make a graph of the number of invasive plant species per class
```{r}
temp6 <- ddply(species_list_2p, .(firstyear, class), nrow)
species_list_3p <- merge(species_list_2p, temp6, by=c("firstyear","class"))

species_list_3p <- subset(species_list_3p, select = -c(number2))
names(species_list_3p)[names(species_list_3p)=="V1"] <- "number3"


print(plot <- ggplot(species_list_3p, aes(x = firstyear, y=number3)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  facet_wrap(~class, nrow=6, scales = "free_y")+
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("Number") + ggtitle("Number of invasive plant species per year per class") +
  scale_y_continuous(limit=c(0,NA),oob=squish))
```

Only make graphs of the number of invasive animal species per class for relevant classes
```{r}
species_list_4a <- subset (species_list_3a, species_list_3a$class %in% c("Actinopterygii", "Amphibia", "Arachnida", "Aves", "Bivalvia", "Gastropoda", "Hydrozoa", "Insecta", "Malacostrata", "Mammalia", "Maxillopoda", "Reptilia"))

print(plot <- ggplot(species_list_4a, aes(x = firstyear, y=number3)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  facet_wrap(~class, nrow=5, scales = "free_y")+
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("Number") + ggtitle("Number of invasive animal species per year per class") +
  scale_y_continuous(limit=c(0,NA),oob=squish))
```


Only make graphs of the number of invasive plant species per class for relevant classes
```{r}
species_list_4p <- subset (species_list_3p, species_list_3p$class %in% c("Liliopsida", "Magnoliopsida", "Pinopsida", "Polypodiopsida"))


print(plot <- ggplot(species_list_4p, aes(x = firstyear, y=number3)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  facet_wrap(~class, nrow=2, scales = "free_y")+
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("Number") + ggtitle("Number of invasive plant species per year per class") +
  scale_y_continuous(limit=c(0,NA),oob=squish))
```

Put these plots in one figure:
1. Set up the multiplot function
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=2, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(c(1,1,2,2))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

2. Make the plots
```{r}
species_list4 <- subset (species_list2, species_list2$class %in% c("Actinopterygii", "Amphibia", "Arachnida", "Aves", "Bivalvia", "Gastropoda", "Hydrozoa", "Insecta", "Malacostrata", "Mammalia", "Maxillopoda", "Reptilia", "Liliopsida", "Magnoliopsida", "Pinopsida", "Polypodiopsida"))

d1 <- ggplot(species_list2, aes(x = firstyear, y=number)) +
  geom_point(colour="black", fill="black", size = 4) +
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") + ggtitle("First record rate of invasive species per year") +
  scale_x_continuous(breaks = seq(1940, 2017, 10))+
  scale_y_continuous(limit=c(0,NA),oob=squish)

d2 <-ggplot(species_list4, aes(x = firstyear, y=number)) +
  geom_point(colour="black", fill="black", size = 4, position="dodge",stat="identity") +
  facet_wrap(~class, nrow=3, scales = "free_y")+
  theme_bw()+
  theme(strip.background =element_rect(fill=NA))+
  theme(strip.text = element_text(colour = "black", face = "bold"))+
  stat_smooth(color="red", fill = NA)+
  xlab("Year") + ylab("First record rate (per year)") +
  scale_y_continuous(limit=c(0,NA),oob=squish)

multiplot(d1, d2)

png(filename = "First record rate of invasive species per year per class.png", width = 1000, height = 900, res=80)
multiplot(a1, a2, a3)
dev.off()

```


Make this figure of the first record rate per five years:
```{r}
table(species_list$class)
```




